---
title: "R Notebook"
output: html_notebook
---

```{r}
Sept11.data <- read.csv("D://1 KSU//MSBA//ST in Business//Assignment//Sept11Travel.csv")
View(Sept11.data)
library(ISLR)
library(forecast)
```

#1. Plot the pre-event AIR time series. What time series components appear?
  Ans: Trend and seasonality appear from the plot.
```{r}
#Plot pre-event Air
Sept11Air.ts <- ts(Sept11.data$Air.RPM..000s., start = c(1990, 1), end = c(2001, 8), freq = 12)
plot(Sept11Air.ts, xlab = "Time", ylab = "Air RPM (000s)", ylim = c(20000000, 70000000))
```

#2. The following figure shows a time plot of the seasonally adjusted pre-September-11 AIR series. Which of the following methods would be adequate for forecasting the series shown in the figure?
   Ans: Linear regression model with trend.
  
 

#3. Specify a linear regression model for the AIR series that would produce a seasonally adjusted series similar to the one shown in the figure below, with multiplicative seasonality. What is the outcome variable? What are the predictors? 
   Ans: The outcome variable are trend and noise, and the predictors will be time.
  
#4. Run the regression model from (3). Remember to use only pre-event data.
  1.What can we learn from the statistical insignificance of the coefficients for October and September?
   Ans: The coefficients of september and october are very close to the regression line which means they do not show a lot of information because it is already given by regression line.
  2.The actual value of AIR (air revenue passenger miles) in January 1990 was 35.153577 billion. What is the residual for this month, using the regression model? Report the residual in terms of air revenue passenger miles.
   Ans: In January 1990, the residual is equal to |(Original Value) - (Valus by Regression Line)| = |-39.1 + 35.15| = 3.95 air revenue passenger miles
```{r}
#3&4:
Sept11Air.ts <- ts(Sept11.data$Air.RPM..000s., 
    start = c(1990, 1), end = c(2004, 3), freq = 12)
library(seasonal)
Z=seas(Sept11Air.ts)
final(Z)
plot(Z)
```
  
#5. Create an ACF (autocorrelation) plot of the regression residuals.
  1.What does the ACF plot tell us about the regression model's forecasts?
  2.How can this information be used to improve the model?
   Ans: The ACF is stationary and exponentially decay, and they are steady time series. If there is any oscillation found in the graph, that indicates the correlation between two seasons.  
```{r}
#5:
Sept11Air.ts.season<- tslm(Sept11Air.ts~ trend+season)
summary(Sept11Air.ts.season)
Acf(Sept11Air.ts.season$residuals, lag.max = 12, main="")
#Let us also display partial autocorrelation
Pacf(Sept11Air.ts.season$residuals, lag.max = 12, main="")
```
#6. Fit linear regression models to Air, Rail, and to Auto with additive seasonality and an appropriate trend. For Air and Rail, fit a linear trend. For Rail, use a quadratic trend. Remember to use only pre-event data. Once the models are estimated, use them to forecast each of the three post-event series.
  1.For each series (Air, Rail, Auto), plot the complete pre-event and post-event actual series overlayed with the predicted series.
  2.What can be said about the effect of the September 11 terrorist attack on the three modes of transportation? Discuss the magnitude of the effect, its time span, and any other relevant aspects.
```{r}
#6
# Plot for Air
Sept11Air.ts <- ts(Sept11.data$Air.RPM..000s., 
                start = c(1990, 1), end = c(2004, 3), freq = 12)
nValid <-32
nTrain <- length(sept11.ts) - nValid

Pre911Air.ts <- window(Sept11Air.ts, start = c(1991, 1), end = c(2001,8))
Post911Air.ts <- window(airdata.ts, start = c(1991, nTrain + 1),
                        end = c(1991, nTrain + nValid))

naive.pred <- forecast::naive(Pre911Air.ts, h = nValid)
snaive.pred <- forecast::snaive(Pre911Air.ts, h = nValid)

plot(sept11.ts,  ylab = "Air Travel", xlab = "Year", bty = "l", 
     xaxt = "n", main = "") 
axis(1, at = seq(1990, 2004, 4), labels = format(seq(1990, 2004, 4)))
lines(naive.pred$mean, lwd = 2, col = "blue", lty = 1)
lines(snaive.pred$mean, lwd = 2, col = "red", lty = 1)

lines(c(2001.75 - 3, 2004.75 - 3), c(0, 3500)) 
lines(c(2001.75, 2001.75), c(0, 3500))
text(1990.25, 2500, "Training")
text(2001.75, 2500, "Validation")
text(2004.25, 2500, "Future")


# Plot for Rail
Sept11Rail.ts <- ts(Sept11.data$Rail.PM, 
                 start = c(1990, 1), end = c(2004, 3), freq = 12)
nValid <-32
nTrain <- length(Sept11Rail.ts) - nValid

Pre911Rail.ts <- window(Sept11Rail.ts, start = c(1991, 1), end = c(2001,8))
Post911Rail.ts <- window(Sept11Rail.ts, start = c(1991, nTrain + 1),
                         end = c(1991, nTrain + nValid))

naive.pred <- forecast::naive(Pre911Rail.ts, h = nValid)
snaive.pred <- forecast::snaive(Pre911Rail.ts, h = nValid)

plot(sept11T.ts,  ylab = "Rail Travel", xlab = "Year", bty = "l", 
     xaxt = "n", main = "") 
axis(1, at = seq(1990, 2004, 4), labels = format(seq(1990, 2004, 4)))
lines(naive.pred$mean, lwd = 2, col = "blue", lty = 1)
lines(snaive.pred$mean, lwd = 2, col = "red", lty = 1)

lines(c(2001.75 - 3, 2004.75 - 3), c(0, 3500)) 
lines(c(2001.75, 2001.75), c(0, 3500))
text(1990.25, 2500, "Training")
text(2001.75, 2500, "Validation")
text(2004.25, 2500, "Future")


#Plot for car
Sept11Car.ts <- ts(Sept11.data$VMT..billions., 
                 start = c(1990, 1), end = c(2004, 3), freq = 12)
nValid <-32
nTrain <- length(Sept11Car.ts) - nValid

Pre911Car.ts <- window(Sept11Car.ts, start = c(1991, 1), end = c(2001,8))
Post911Car.ts <- window(Sept11Car.ts, start = c(1991, nTrain + 1),
                         end = c(1991, nTrain + nValid))

naive.pred <- forecast::naive(Pre911Car.ts, h = nValid)
snaive.pred <- forecast::snaive(Pre911Car.ts, h = nValid)

plot(sept11c.ts,  ylab = "Car Travel", xlab = "Year", bty = "l", 
     xaxt = "n", main = "") 
axis(1, at = seq(1990, 2004, 4), labels = format(seq(1990, 2004, 4)))
lines(naive.pred$mean, lwd = 2, col = "blue", lty = 1)
lines(snaive.pred$mean, lwd = 2, col = "red", lty = 1)
lines(c(2001.75 - 3, 2004.75 - 3), c(0, 3500)) 
lines(c(2001.75, 2001.75), c(0, 3500))

text(1990.25, 2500, "Training")
text(2001.75, 2500, "Validation")
text(2004.25, 2500, "Future")
```

  
  
